---
title: "Experiments with rayrender"
---

## Rayrender and cylinders

After playing around with the system [corpuscular](https://github.com/paezha/corpuscular) I started thinking of monumental sculptural spaces. Here I experiment with creating vast vistas with columns (pilares) as seen from various vantage points.

These packages are needed:
```{r}
library(dplyr)
library(glue)
library(MetBrewer)
library(MexBrewer)
library(rayrender)
```

Sample a number to use as a random seed:
```{r}
#seed <- 2687084
seed <- sample.int(10000000, 1)
```

Randomly choose a color palette from {MetBrewer} or {MexBrewer}:
```{r}
# Select collection of color palettes
edition <- sample(c("MexBrewer", "MetBrewer"), 1)

if(edition=="MexBrewer"){
  # Randomly select a color palette (MexBrewer Edition)
  palette_name <- sample(c("Alacena", "Atentado", "Aurora", "Concha", "Frida", "Revolucion", "Ronda", "Tierra"), 1)
  # col_palette <- mex.brewer(palette_name, n = 25)
  col_palette <- mex.brewer(palette_name, n = 15)
  
}else{
  # Randomly select a color palette (MetBrewer Edition)
  palette_name <- sample(c("Archaumbault", "Austria", "Benedictus", "Cassatt1", "Cassatt2", "Cross", "Degas", "Demuth", "Derain", "Egypt", "Gauguin", "Greek", "Hiroshige", "Hokusai1", "Hokusai2", "Hokusai3", "Homer1", "Homer2", "Ingres", "Isfahan1", "Isfahan2", "Java", "Johnson", "Juarez", "Kandinsky", "Klimt", "Lakota", "Manet", "Monet", "Moreau", "Morgenstern", "Nattier", "Navajo", "NewKingdom", "Nizami", "OKeeffe1", "OKeeffe2", "Paquin", "Peru1", "Peru2", "Pillement", "Pissaro", "Redon", "Renoir", "Robert", "Signac", "Tam", "Tara", "Thomas", "Tiepolo", "Troy", "Tsimshian", "VanGogh1", "VanGogh2", 'VanGogh3', "Veronese", "Wissing"), 1)
  col_palette <- met.brewer(palette_name, n = 15)
}
```

Create a data frame with the parameters for the columns. The position in x is on the plane (left-right), y is the vertical axis, and z is on the plane (front-back). The radius must be positive. Given the spacing between columns, the radius should not exceed perhaps 1/8 or so of the spacing. Use the color palette chosen:
```{r}
df <- expand.grid(x = c(-3, 3), # This is to create two parallel rows of columns; consider adding more
                  z = seq(20, -50, -2.5)) |> # This is the spacing between columns in the z axis (front-back)
  mutate(y = 0, # The columns are on the ground, not floating, not sunk
         l = runif(1, min= 100, max = 300), # The columns are _tall_, between 100 and 300 units tall
         r = 0.15, # The radius is a fraction of the smallest spacing of the columns (which is at the moment 2 in x)
         color = sample(col_palette, 
                                n(), 
                                replace = TRUE))
```

To create the columns for the scene, initialize an empty data frame and then use `rayrender::cylinder()`. Experiment with different materials (`diffuse()`, `glossy()`, etc.). Populate the parameters to create the sphere:
```{r}
obj <- data.frame()

for(i in 1:nrow(df)){
  obj <- rbind(obj,
               cylinder(x = df$x[i],
                      y = df$y[i],
                      z = df$z[i],
                      radius = df$r[i],
                      l = df$l[i],
                      material = diffuse(color=df$color[i])))
}
```

Initialize a scene with a white ground:
```{r}
set.seed(seed)

bkg_c <- sample.int(12, 1) + 3
scene <- generate_ground(material = diffuse(color = col_palette[bkg_c]))
```

Add the objects to the scene, including a source of light with random intensity:
```{r}
set.seed(seed)

scene <- scene |> 
  add_object(objects = obj) |>
  #light source
   add_object(sphere(x = sample(c(-30, 30), 1),
                    y = 20,
                    z = -60,
                    r = 4,
                    material=light(intensity = rlnorm(1, meanlog = 4),
                                   invisible = TRUE)))
```

Render the scene using randomly drawn colors from the chosen palette for the background. 
```{r}
set.seed(seed)

x_from <- runif(1, min = -3, max = 3)
z_from <- runif(1, min = 25, max = sqrt(30^2 - x_from^2))
y_from <- runif(1, min = 0, max = 80) #runif(1, min = 0, max = sqrt(30^2 - x_from^2 - y_from^2))

render_scene(file = glue::glue("pilares-{seed}.png"),
             scene, 
             parallel = TRUE,
             ambient_light = TRUE,
             width = 1500, 
             height = 1500, 
             samples = 150 + abs(rnorm(1, 0, 100)), 
             backgroundhigh = col_palette[bkg_c],
             backgroundlow = col_palette[bkg_c - sample.int(4, 1)],
             lookfrom = c(x_from, z_from, y_from),#c(0, rlnorm(1, 1, 2), runif(1, min = 15, max = 30)),
             lookat = c(0, 2, 1))
```

